---
title: "Tall Grass and Weed Statistics"
author: "Harrisonburg"
date: "Sat Sep  3 05:08:12 2022"
output: 
 html_document:
  theme: cosmo

---


## Code enforcement complaints mapped.
```{r, "libraries",  echo=FALSE, message=FALSE }
##https://yihui.org/knitr/options/#code-evaluation
# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
names(cbbPalette) <- c("black","orange","sky","green","yellow","blue","red","pink")
##https://github.com/walkerke/umich-workshop-2022

## https://permits.harrisonburgva.gov/
## https://rpubs.com/nickbearman/r-google-map-making
## https://ajsmit.github.io/Intro_R_Official/mapping-google.html

## https://www.gpsvisualizer.com/geocoding.html
## https://geocoding.geo.census.gov/geocoder/locations/onelineaddress?address=560%20Norwood%20St%20Harrisonburg%20Virginia&benchmark=4

#https://stackoverflow.com/questions/4090169/elegant-way-to-check-for-missing-packages-and-install-them

r = getOption("repos") 
r["CRAN"] = "http://lib.stat.cmu.edu/R/CRAN/" 
options(repos = r)
rm(r)

using <- function(...) {
    libs<-unlist(list(...))
    req<-unlist(lapply(libs,require,character.only=TRUE))
    need<-libs[req==FALSE]
    if(length(need)>0){ 
        install.packages(need)
        lapply(need,require,character.only=TRUE)
    }
}

libs <- c("Hmisc","tidyverse","dplyr","clock","tigris","ClusterR","cluster","ggmap","ggplot2","rvest","showtext","sp","sf","tidycensus","tmap","icons","fontawesome")
using(libs)
```



```{r, setup,  echo=FALSE, message=FALSE, cache=TRUE }


if(system("whoami", intern=TRUE)=="lchen")
    setwd("/Users/lchen/shared/yard") else{
                                          system("mount_smbfs lchen@172.16.0.7/share /Volumes/share")
                                          setwd("/Volumes/shared/yard")}
to418 <- read.csv("tgw/2022/exports/to418.csv")
to1018 <- read.csv("tgw/2022/exports/to1018.csv")
to1619 <- read.csv("tgw/2022/exports/to1619.csv")
to919 <- read.csv("tgw/2022/exports/to919.csv")
to919 <- read.csv("tgw/2022/exports/to919.csv")
to520 <- read.csv("tgw/2022/exports/to520.csv")
to621 <- read.csv("tgw/2022/exports/to621.csv")
to1121 <- read.csv("tgw/2022/exports/to1121.csv")
to522 <- read.csv("tgw/2022/exports/to522.csv")
to722 <- read.csv("tgw/2022/exports/to722.csv")
to922 <- read.csv( "tgw/2022/exports/to922.csv")
to1222 <- read.csv("tgw/2022/exports/to1222.csv")
code <- rbind(to418,
      to1018,
      to1619,
      to919,
      to919,
      to520,
      to621,
      to1121,
      to522,
      to722,
      to922,
      to1222)
code <- code[!duplicated(code),]

drops <- c("IsDeleted","TotalFees","UnpaidFees","Value","CouncilDistrict","Next.Compliance.Date","CanPayFees","UserFavID","UserID","FavType","FavID","IsActive","IsDeleted1","ContactID","CanEdit","CanView","CanRenew","NeedAccept","Next.Hearing.Date","Last.Hearing.Date","LU_JurisdictionID","IsPublic","PubicDocCount","IsLockedExt", "IsLockedInt","Fav")
code <- code[, !(names(code) %in% drops)]
code[code$Location=="1552 Old Richmond Cir","Location"] <- "1252 Old Richmond Cir" 
code[code$Location=="1556 Old Richmond Cir","Location"] <- "1256 Old Richmond Cir"
code[code$Location=="0 Star Crest Dr","Location"] <- "1080 Starcrest Dr"
code[code$Location=="901 S High Street","Location"] <- "205 Sunrise Ave"
code[code$Location=="205 Sunrise Ave", c("YWGS84","XWGS84")] <- c(38.441041699270514, -78.88489286387966)
code[code$Location=="66 Pleasant Hill rd, 60 Pleasant Hill Rd","Location"] <- "66 Pleasant Hill Rd"
code[code$Location=="66 Pleasant Hill Rd", c("YWGS84","XWGS84")] <- c(38.428048249736136, -78.89348158194454)
code[code$Location=="1083 Lois Lane", c("YWGS84","XWGS84")] <- c(38.4165629683439, -78.87282929749068)
code[code$Location=="371 W Water St -Apt B", c("YWGS84","XWGS84")] <- c(38.449331296072266 ,-78.87555135613829)
code[1129,c("YWGS84","XWGS84")] <- c(38.445681504207094, -78.87579007681562)

missings <- unique(c(code[code$XWGS84> -20|code$XWGS84< -1000, "Location" ],code[code$YWGS84> 2000|code$YWGS84< 30, "Location"]))

code[code$Location==missings[1]  ,c("XWGS84")] <- -78.86329722699998
code[code$Location==missings[2]  ,c("XWGS84")] <- -78.85689034099994
code[code$Location==missings[3]  ,c("XWGS84")] <- -78.86860535299996
code[code$Location==missings[4]  ,c("XWGS84")] <- -78.89740181099995
code[code$Location==missings[5]  ,c("XWGS84")] <- -78.85792131799997
code[code$Location==missings[6]  ,c("XWGS84")] <- -78.85720977999995
code[code$Location==missings[7]  ,c("XWGS84")] <- -78.86409398899997
code[code$Location==missings[8]  ,c("XWGS84")] <- -78.85405233399996
code[code$Location==missings[9]  ,c("XWGS84")] <- -78.87034136799997
code[code$Location==missings[10]  ,c("XWGS84")] <- -78.84623136199997
code[code$Location==missings[11]  ,c("XWGS84")] <- -78.85982703299999
code[code$Location==missings[12]  ,c("XWGS84")] <- -78.85703166499997
code[code$Location==missings[13]  ,c("XWGS84")] <- -78.85408945199998
code[code$Location==missings[14]  ,c("XWGS84")] <- -78.86749465599996
code[code$Location==missings[15]  ,c("XWGS84")] <- -78.86590031799994
code[code$Location==missings[16]  ,c("XWGS84")] <- -78.86904349499997
code[code$Location==missings[17]  ,c("XWGS84")] <- -78.86708798299998
code[code$Location==missings[18]  ,c("XWGS84")] <- -78.84802488399998
code[code$Location==missings[19]  ,c("XWGS84")] <- -78.88383759299995
code[code$Location==missings[20]  ,c("XWGS84")] <- -78.85106121699994
code[code$Location==missings[21]  ,c("XWGS84")] <- -78.85103213899998
code[code$Location==missings[22]  ,c("XWGS84")] <- -78.85100305999998
code[code$Location==missings[23]  ,c("XWGS84")] <- -78.85097249099994
code[code$Location==missings[1]  ,c("YWGS84")] <- 38.45395613300008
code[code$Location==missings[2]  ,c("YWGS84")] <- 38.44490151800005
code[code$Location==missings[3]  ,c("YWGS84")] <- 38.44351018100008
code[code$Location==missings[4]  ,c("YWGS84")] <- 38.43486111000004
code[code$Location==missings[5]  ,c("YWGS84")] <- 38.44475642100008
code[code$Location==missings[6]  ,c("YWGS84")] <- 38.44486554200006
code[code$Location==missings[7]  ,c("YWGS84")] <- 38.44107083600005
code[code$Location==missings[8]  ,c("YWGS84")] <- 38.43851355400005
code[code$Location==missings[9]  ,c("YWGS84")] <- 38.45439405000007
code[code$Location==missings[10]  ,c("YWGS84")] <-38.43408279400006
code[code$Location==missings[11]  ,c("YWGS84")] <-38.44930966600003
code[code$Location==missings[12]  ,c("YWGS84")] <-38.44295990600005
code[code$Location==missings[13]  ,c("YWGS84")] <-38.42154672500004
code[code$Location==missings[14]  ,c("YWGS84")] <-38.44512364500008
code[code$Location==missings[15]  ,c("YWGS84")] <-38.45973579700006
code[code$Location==missings[16]  ,c("YWGS84")] <-38.44763803900002
code[code$Location==missings[17]  ,c("YWGS84")] <-38.44506554800006
code[code$Location==missings[18]  ,c("YWGS84")] <-38.44624103700005
code[code$Location==missings[19]  ,c("YWGS84")] <-38.45194733000005
code[code$Location==missings[20]  ,c("YWGS84")] <-38.45884121100005
code[code$Location==missings[21]  ,c("YWGS84")] <-38.45880061000002
code[code$Location==missings[22]  ,c("YWGS84")] <-38.45876001000005
code[code$Location==missings[23]  ,c("YWGS84")] <-38.45872046200003
code[code$Status=="","Status"] <- "No Status"
code[code$Status=="2nd Notice - Tall Weeds and Grass", "Status"] <- "2nd Notice - TGW"
code[code$Guid== "a4571e39-ab32-4e00-908b-da17f1237d60","Status"] <- "No Violations Found"


privacy <-c(1799,1794,1740,1741,1731,1718,1717,1692,1651,1634,1611,1593,1584,1583,1580,1560,1540,1537,1526,1523,1501,1431,1317,1258,1237,1195,1193,1189,1181,1180,1175,1167,1162,1149,1150,1143,1140,1139,1123,1107,1084,1082,1077,1063,1036,1030,1029,1026,1022,1018,1017,1016,1015,1014,998,993,989,975,969,965,957,955,951,935,922,917,916,892,893,884,883,877,871,870,869,867,864,862,860,859,848,845,841,819,762,728,724,718,705,704,703,701,693,692,668,666,636,630,628,584,551,531,487,467,465,455,450,407,391,390,375,287,281,236,234,224,211,199,196,176,166,132,122,47,46,27,7,1797,1775)

#tenants <- code %>% grep("tenant", "Descr", ignore.case= TRUE, value=FALSE) 
tenants <- grep("tenant",code$Descr, ignore.case=TRUE, value=FALSE)
privacy <- unique(c(tenants,privacy))
code[privacy,"Descr"] <- unlist(lapply(code[privacy,"Descr"], gsub, pattern="[0-9]+?", replacement="x"))
code <- code %>% filter( Guid!="661f722c-936e-4bd1-a974-eb37663b0a45"& Guid!="11b32ee3-b20f-49fa-87ac-621f9e00ab30")
```
```{r, early tgw, echo=FALSE, message=FALSE, cache=FALSE}

## if(system("whoami", intern=TRUE)=="lchen")
##     setwd("/Users/lchen/Sites/yard") else{
##                                           system("mount_smbfs lchen@172.16.0.7/share /Volumes/share")
##                                           setwd("/Volumes/shared/Sites/yard")}

tgwraw13 <- read.csv("tgwraw13.csv")
geocoded12 <- read.csv("addresses.csv")
geocoded13 <- read.csv("addresses13.csv")
tgwfull <- cbind(tgwraw13,geocoded13)
tgwfull <- tgwfull[tgwfull$Source=="citizen concern",]
tgw12 <- read.csv("tgw2.csv")
tgw12 <- cbind(tgw12,"unknown")
colnames(tgw12)[dim(tgw12)[2]] <- "Source.of.Concern"
levels(tgw12$Source.of.Concern) <- c("unknown","proactive")
tgw12$Source.of.Concern[c(130:133,122,120,114,115)] <- "proactive"
tgwdate12 <- as.Date(tgw12$Date.Received)
tgwdate <- as.Date(tgwfull$Date.Received,"%m-%d-%Y")

tgw12$X <- tgwdate12
colnames(tgw12)[1] <- "tgwdate"
tgw12 <- cbind(tgw12,tgw12$Comments)
colnames(tgw12)[19] <- "tgwfull$Comments"
###-
earlytgw <- cbind(tgwdate,tgwfull,tgwfull$Comments)

## earlytgw <- rbind(tgw12[,c(colnames(tgw12)[-match("Source.of.Concern",colnames(tgw12))],"Source.of.Concern")],

##                   earlytgw[,c(colnames(earlytgw)[-match("Source.of.Concern",colnames(earlytgw))],"Source.of.Concern")])
earlytgw <- bind_rows(tgw12,earlytgw)

names(earlytgw)[match("tgwfull$Comments",colnames(earlytgw))] <- "type"
names(earlytgw)[match("Violation.Status.Open.Closed",colnames(earlytgw))] <- "Status"
##earlytgw <- bind_rows(earlytgw,tgw12)
earlytgw$TypeName <- "Tall Grass and Weeds"
earlytgw$type <- earlytgw$Comments
earlytgw$type <- as.factor(earlytgw$type)
levels(earlytgw$type)[grep("[mb]ul[tp][it][pi]le", levels(earlytgw$type))] <- "group"
levels(earlytgw$type)[grep("unk", levels(earlytgw$type))] <- "junk"
levels(earlytgw$type)[grep("inop", levels(earlytgw$type))] <- "junk"
levels(earlytgw$type)[grep("oreclo", levels(earlytgw$type))] <- "foreclosure"
levels(earlytgw$type)[grep("bank owned", levels(earlytgw$type))] <- "foreclosure"
levels(earlytgw$type)[grep("asset company", levels(earlytgw$type))] <- "foreclosure"
levels(earlytgw$type)[grep("garden", levels(earlytgw$type))] <- "garden"
levels(earlytgw$type)[grep("back",levels(earlytgw$type))] <- "bounced"
levels(earlytgw$type)[grep("trouble finding",levels(earlytgw$type))] <- "bounced"
levels(earlytgw$type)[grep("[tT]all", levels(earlytgw$type))] <- "TGW"
levels(earlytgw$type)[grep("but", levels(earlytgw$type))] <- "TGW"
levels(earlytgw$type)[grep("RGR", levels(earlytgw$type))] <- "mowed"
levels(earlytgw$type)[c(24,10,9)] <- "mowed"
levels(earlytgw$type)[c(16,13)] <- "common"
levels(earlytgw$type)[c(18:21,15,16,10:12,6:8,4)] <- "TGW"


earlytgw$Status[grep("[0-9]", earlytgw$Status)] <- "No Status"
earlytgw$Status[grep("Closed", earlytgw$Status,ignore.case=TRUE)] <- "No Status" 
earlytgw$Status[grep("n-a", earlytgw$Status)] <- "No Status"
earlytgw$Status[earlytgw$Status=="no violation"] <- "No Violations Found"
earlytgw$Status[earlytgw$Status=="No violation"] <- "No Violations Found"
earlytgw$Status[earlytgw$Status=="unfounded"] <- "No Violations Found"
earlytgw$Status[earlytgw$Status==""] <- "No Status"
earlytgw$Status[earlytgw$Status=="Rosalyn"] <- "No Status"
earlytgw$Status[earlytgw$Status=="Generate Sign Civil Penalty Notice"] <- "Sign Notice"



tgwdate <- c(tgwdate12,tgwdate)[earlytgw$Source.of.Concern=="unknown"|earlytgw$Source.of.Concern=="citizen concern"]
earlytgw <- earlytgw[earlytgw$Source.of.Concern=="unknown"|earlytgw$Source.of.Concern=="citizen concern",]
breaksv <- as.Date(c("2011-04-01","2011-7-24","2011-09-01","2012-03-01","2012-07-24","2012-09-01","2012-12-1","2013-03-01","2013-07-24","2013-12-1"))
era <- cut(unclass(tgwdate),unclass(breaksv),lab=FALSE)
era <- as.factor(era)
levels(era) <- c("2011pre","2011post","2012pre","2012post","2012xtended","2013pre","2013post","2013extended")
era <- factor(era,levels=c("2011pre","2012pre","2013pre","2011post","2012post","2013post","2012xtended","2013extended"))
earlytgw <- cbind(earlytgw,era)
code$FilingDate <- as.Date(code$FilingDate, format="%m/%d/%Y %r")
names(earlytgw)[names(earlytgw)=="tgwdate"] <- "FilingDate"
earlytgw$FilingDate <- as.Date(earlytgw$FilingDate, format="%Y-%m-%d")
names(earlytgw)[names(earlytgw)=="latitude"] <- "YWGS84"
names(earlytgw)[names(earlytgw)=="longitude"] <- "XWGS84"
ocode <- code[code$XWGS84!=-1,]#come back to this.
```
```{r, demographicfs, eval=TRUE, echo=FALSE, message=FALSE}
if(system("whoami", intern=TRUE)=="lchen"){tigris_cache_dir("/Users/lchen/shared/yard/tigris/")} else{
 tigris_cache_dir("/Volumes/shared/yard/tigris/") }
readRenviron('~/.Renviron')
options(tigris_use_cache = TRUE)

burglatino2020 <- get_decennial(year=2020, geography="block", county="Harrisonburg", state="VA", variables=c("P2_002N","P2_003N","P2_004N","P2_005N","P2_006N","P2_007N","P2_008N","P2_009N"), summary_var = "P2_001N",geometry=TRUE, cache_table=TRUE) %>% mutate(percent=100*(value/summary_value))

burgraceall2020 <- get_decennial(year=2020, geography="block", county="Harrisonburg", state="VA", variables=c("P1_002N","P1_003N","P1_004N","P1_005N","P1_006N","P1_007N","P1_008N","P1_009N"),summary_var="P1_001N",geometry=TRUE, cache_table=TRUE)

burgraceall2020 <- burgraceall2020 %>% mutate(percent=100*(value/summary_value))
tract_grass_sf <- filter(burgraceall2020, variable=="P1_003N")%>% mutate(wite=value-filter(burglatino2020, variable=="P2_005N")$value) %>% mutate(percent=100*(value/summary_value)) #white

```

```{r, spatial, echo=FALSE, message=FALSE}
 
if(system("whoami", intern=TRUE)=="lchen")
    setwd("/Users/lchen/shared/yard") else{
                                          system("mount_smbfs lchen@172.16.0.7/share /Volumes/share")
                                          setwd("/Volumes/shared/yard")}

crs_utm <- st_crs(burgraceall2020)

code_sp <- st_as_sf(code, coords = c("XWGS84", "YWGS84"), crs = crs_utm )
earlytgw_sp <- st_as_sf(earlytgw, coords = c("XWGS84", "YWGS84"), crs = crs_utm )
##https://jonathanduggins.com/data-matters (R slides)
code_pz_sp <- filter(code_sp,Division=="P&Z") 
#https://r-spatial.github.io/sf/articles/sf5.html
code_pz_nuisance_sp <- filter(code_pz_sp, TypeName %in% c("Junk, Trash and Debris","Tall Grass and Weeds","Inoperable Vehicle","Signage"))
code_pz_nuisance_sp <- bind_rows(code_pz_nuisance_sp, earlytgw_sp)
code_pz_nuisance_sp <- within(code_pz_nuisance_sp, TypeName <-factor(TypeName, levels=names(sort(table(TypeName),decreasing=TRUE))))
code_pz_nuisance_sp <- within(code_pz_nuisance_sp, Status <-factor(Status, levels=names(sort(table(Status),decreasing=TRUE))))

```

```{r, icons, echo=FALSE, message=FALSE }

fontawesome::fa_png("building-wheat", fill=cbbPalette["blue"])
system("mv building-wheat.png building-none.png")
fontawesome::fa_png("car-side", fill="grey")
fontawesome::fa_png("building-wheat")
fontawesome::fa_png("trash", fill=cbbPalette["yellow"])
fontawesome::fa_png("trash-can", fill=cbbPalette["yellow"])
fontawesome::fa_png("couch", fill=cbbPalette["red"])
building_wheat <- tmap_icons("building-wheat.png", height=12)
building_none <- tmap_icons("building-none.png", height=12)
car_side <- tmap_icons("car-side.png", height=12)
trash <- tmap_icons("trash.png", height=12)
trash_can <- tmap_icons("trash-can.png", height=12)
couch <- tmap_icons("couch.png", height=12)

```
```{r, violation partitioning,  echo=FALSE, message=FALSE}
##code_pz_nuisance_sp <- code_pz_nuisance_sp %>% filter(FilingDate > as.Date("2021-9-12"))
grass <- code_pz_nuisance_sp %>% filter(TypeName=="Tall Grass and Weeds"&!Status%in% c("No Violations Found","No Status","Inspection"))

nograss <- code_pz_nuisance_sp %>% filter(TypeName=="Tall Grass and Weeds"&Status%in% c("No Violations Found","No Status","Inspection"))

junk <- code_pz_nuisance_sp %>% filter(TypeName=="Junk, Trash and Debris")
couches <- junk[unlist(lapply(junk$Descr, str_detect, "ouch|urni|hair")),]

jnk <- junk[!unlist(lapply(junk$Descr, str_detect, "ouch|urni|hair")),]
nojnk <- jnk %>% filter(Status%in%c("No Violations Found","No Status","Inspection"))
jnk <- jnk %>% filter(!Status%in%c("No Violations Found","No Status","Inspection"))
cars <- code_pz_nuisance_sp %>% filter(TypeName=="Inoperable Vehicle")

```

```{r, 'main map setup', eval = TRUE, echo=FALSE, message=FALSE}
tmap_mode("view")
demographic <- tm_shape(tract_grass_sf, name="Percent Non-Latino White") +
  tm_fill(
    col = "percent",
    palette = "-Greens",
    style = "cont",
    contrast = c(0.1, 1),
    title = "Percent Non-Latino White",
    id = "summary_value", #"GEOID",
    popup.vars = c("percent","value","summary_value"),
    showNA = FALSE,
    alpha = 0.8) +
    tm_borders(col = "darkgray", lwd = 0.7)

```
```{r, 'main map', echo=FALSE, message=FALSE , eval=TRUE}
## fontawesome::fa_png("trash-can", fill=cbbPalette["yellow"])
## to do as grassdot <- as_dot_density(value=) follow https://youtu.be/KMOTZij9qhU?t=9770 and tgw.R for tabling and left joining counts to polygons. 

###tmap_mode("view")

grassmap <- demographic +
    tm_shape(grass, name="Tall Grass and Weed")+
    tm_symbols(shape=building_wheat, id = "Status",
               popup.vars = c("Origin") , size=.5, jitter=.3, alpha=.5)+
    tm_shape(nograss, name="Unfounded Tall Grass and Weed")+
    tm_symbols(shape=building_none, id = "Status",
             popup.vars = c("Descr","Origin") ,size=.5, jitter=.3, alpha=.5) +
    tm_shape(cars, name="Inoperable Vehicle")+
    tm_symbols(shape=car_side, id = "Origin",
             popup.vars = c("Descr","Status") , size=.5, jitter=.3) +
    tm_shape(jnk, name="Junk, Trash and Debris")  +
    tm_symbols( shape=trash, id = "Origin",
             popup.vars = c("Status"), size=.5, jitter=.3) +
    tm_shape(nojnk, name="Unfounded Junk, Trash and Debris")+
    tm_symbols(shape=trash_can, id = "Origin",
             popup.vars = c("Descr","Status") , size=.5, jitter=.3) +
    tm_shape(couches, name="Couch on Porch") +
    tm_symbols(shape=couch, id = "Origin",
               popup.vars = c("Status"), size=.5, jitter=.3, alpha=.5) + tm_layout("Aesthetic nuisance complaints in Harrisonburg Virginia October 2017 - July 2022 \n jittered with error for privacy." )

grassmap+ggtitle("Harrisonburg Aesthetic Nuisance Complaints", subtitle="Complaints 2011-2013, 2017-2022")
```

```{r, code analysis, echo=FALSE, message=FALSE}
tgw <- code_pz_nuisance_sp

tgw <- within(tgw, TypeName <-factor(TypeName, levels=names(sort(table(TypeName),decreasing=FALSE))))

```
## Code enforcement complaints over time.


```{r widegraphss, message=FALSE, echo=FALSE, fig.width=12, fig.height=4}

seasons <- as.Date(c("2011-4-1","2011-11-1","2012-4-1","2012-11-1","2013-4-1","2013-11-1","2014-4-1","2014-11-1","2015-4-1","2015-11-1","2016-4-1","2016-11-1","2017-4-1","2017-11-1","2018-4-1","2018-11-1","2019-4-1","2019-11-1","2020-4-1","2020-11-1","2021-4-1","2021-11-1","2022-4-1","2022-11-1"))

ggplot(data=tgw, aes(x=FilingDate, y=Status), col=TypeName)+ geom_jitter(alpha=.5, pch=1) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+scale_x_date( minor_breaks="year", date_labels="%y-%m", breaks= seasons)+ggtitle("All Planning and Zoning enforcement status")+xlab("TGW enforcement season")

ggplot(data=tgw, aes(x=as.Date(FilingDate,format="%m/%d/%Y %r"), fill=TypeName))+geom_histogram(binwidth=30, colour="white")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+scale_x_date( minor_breaks="year", date_labels="%y-%m" , breaks= seasons)+ggtitle("All Planning and Zoning enforcement status")+xlab("TGW enforcement season")
```

```{r widegraphss TGW, message=FALSE, echo=FALSE, fig.width=12, fig.height=4}
tgw <- tgw %>% filter(TypeName=="Tall Grass and Weeds")

status_l <- c("No Violations Found", "No Status",  "Inspection", "Verbal Notice", "Tall Weeds and Grass Notice", "Correction Notice", "Return Receipt Rc'd", "Door Hanger", "2nd Notice - TGW", "Violations Remedied")
status_l <- rev(status_l)

tgw <- within(tgw, Status <- factor(Status, levels= status_l ))

tgw %>%ggplot( aes(x=as.Date(FilingDate, format="%m/%d/%Y %r"), y=Status, col=Status))+ geom_jitter(alpha=.5, pch=1) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+scale_x_date( minor_breaks="year", date_labels="%y-%m" ,breaks= seasons)+ggtitle("TGW enforcement status")+xlab("TGW enforcement season")

tgw <- tgw %>% mutate(Status=relevel(Status, 8))

## tgw <- within(tgw, Status <-factor(Status, levels=names(sort(table(Status),decreasing=FALSE))))

ggplot(data=tgw, aes(x=as.Date(FilingDate,format="%m/%d/%Y %r"), fill=Status))+geom_histogram(binwidth=30, colour="white")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+scale_x_date( minor_breaks="year", date_labels="%y-%m" ,breaks=seasons)+ggtitle("TGW enforcement status")+xlab("TGW enforcement season")+geom_vline(xintercept=as.Date("2013-8-13"), col=cbbPalette["red"])+geom_text(label="MLK Renaming",aes(as.Date("2013-8-13"), 40))
```
	
## Sentiments toward natural yards in Harrisonburg

```{r, 'survey', echo=FALSE, message=FALSE}
##setwd("/Volumes/shared/yard/")
nomow <- read.csv("nomow.csv")
nomow$supports[nomow$supports==" yes"] <- "yes"
nomow <- within(nomow, city <-factor(city , levels=names(sort(table(city),decreasing=TRUE))))
nomow[c(73,114),"different"] <- ""
with(nomow,table(supports))
```

On May 7, 2022, [WHSV](https://www.whsv.com/2022/05/06/no-mow-may-puts-lawnmower-pause-help-bees/) aired a story on No Mow May. On their [Facebook](https://www.facebook.com/WHSV.TV/posts/10159997517915336) page, they asked if viewers would consider participating in No Mow May. The `r sum(table(nomow$supports))` responses to the post were coded and analyzed. Notable features are that Harrisonburg overwhelmingly supported No Mow May with support of `r 100*(table(nomow$supports)["yes"]/sum(table(nomow$supports)))`%. Support in Harrisonburg featured acceptance of difference, whereas being different was viewed negatively outside Harrisonburg.

```{r, 'survey analysis', echo=FALSE, message=FALSE}
ggplot(data=nomow,aes(x=city, fill=supports))+ geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("WHSV May 7 Facebook responses on No Mow May preference", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")

ggplot(data=nomow %>% subset(supports=="yes"), aes(x=city, fill=different))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("Difference theme by No Mow May supporters.", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
ggplot(data=nomow %>% subset(supports=="no"), aes(x=city, fill=different))+geom_bar()+theme(axis.text.x = element_text(an gle = 90, vjust = 0.5, hjust=1))+ggtitle("Difference theme by No Mow May opponents.", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")


ggplot(data=nomow %>% subset(supports=="yes"), aes(x=city, fill=HOA))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("HOA theme by No Mow May supporters.", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
ggplot(data=nomow %>% subset(supports=="no"), aes(x=city, fill=HOA))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("HOA theme by No Mow May opponents.", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")



ggplot(data=nomow %>% subset(supports=="yes"), aes(x=city, fill=laws))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("Laws theme by No Mow May supporters", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
ggplot(data=nomow %>% subset(supports=="no"), aes(x=city, fill=laws))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("Laws theme by No Mow May opponents", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")


ggplot(data=nomow %>% subset(supports=="yes"), aes(x=city, fill=mowing))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("Difficult to mow theme by No Mow May supporters", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
ggplot(data=nomow %>% subset(supports=="no"), aes(x=city, fill=mowing))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("Difficult to mow theme by No Mow May opponents", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")


ggplot(data=nomow %>% subset(supports=="yes"), aes(x=city, fill=critters))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("Attract critters theme by No Mow May supporters", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
ggplot(data=nomow %>% subset(supports=="no"), aes(x=city, fill=critters))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("Attract critters theme by No Mow May opponents", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")

ggplot(data=nomow %>% subset(supports=="yes"), aes(x=city, fill=flowers))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("Have flowers theme by No Mow May supporters", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
ggplot(data=nomow %>% subset(supports=="no"), aes(x=city, fill=flowers))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("Have flowers theme by No Mow May opponents", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")


harrisonburg <- nomow %>% subset(city=="Harrisonburg")
ggplot(data=harrisonburg, aes(x=supports, fill=different))+geom_bar() +ggtitle("Difference theme by No Mow May Harrisonburg residents", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
outsiders <- nomow %>% subset(city!="Harrisonburg")
ggplot(data=outsiders, aes(x=supports, fill=different))+geom_bar() +ggtitle("Difference theme by No Mow May non-Harrisonburg residents", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
```
	
