---
title: "Tall Grass and Weed Statistics"
author: "Harrisonburg"
date: "Sat Sep  3 05:08:12 2022"
output: 
 html_document:
  theme: cosmo

---


## Code enforcement complaints mapped.
What follows is an R code snippet that can be stepped through with C-c C-n in emacs, or in the equivalent way in RStudio
The code evaluation instructions in the R chunck headers are explained at https://yihui.org/knitr/options/#code-evaluation
We start with settings and loading libraries. The commented URLs are references for how we did things in the code.

```{r, "libraries",  echo=FALSE, message=FALSE }
## We select an accessible color pallate with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
names(cbbPalette) <- c("black","orange","sky","green","yellow","blue","red","pink")


## https://stackoverflow.com/questions/4090169/elegant-way-to-check-for-missing-packages-and-install-them
r = getOption("repos") 
r["CRAN"] = "http://lib.stat.cmu.edu/R/CRAN/" 
options(repos = r)
rm(r)

using <- function(...) {
    libs<-unlist(list(...))
    req<-unlist(lapply(libs,require,character.only=TRUE))
    need<-libs[req==FALSE]
    if(length(need)>0){ 
        install.packages(need)
        lapply(need,require,character.only=TRUE)
    }
}

libs <- c("Hmisc","tidyverse","dplyr","clock","tigris","ClusterR","cluster","ggmap","ggplot2","rvest","showtext","sp","sf","tidycensus","tmap","icons","fontawesome")
using(libs)
```


Next, we read in and clean the tall grass and weed complaints exported manually from the [Harrisonburg permitting website](https://permits.harrisonburgva.gov/) advanced search interface.

```{r, setup,  echo=FALSE, message=FALSE, cache=TRUE }
## Reading in the files
to418 <- read.csv("tgw/2022/exports/to418.csv")
to1018 <- read.csv("tgw/2022/exports/to1018.csv")
to1619 <- read.csv("tgw/2022/exports/to1619.csv")
to919 <- read.csv("tgw/2022/exports/to919.csv")
to919 <- read.csv("tgw/2022/exports/to919.csv")
to520 <- read.csv("tgw/2022/exports/to520.csv")
to621 <- read.csv("tgw/2022/exports/to621.csv")
to1121 <- read.csv("tgw/2022/exports/to1121.csv")
to522 <- read.csv("tgw/2022/exports/to522.csv")
to722 <- read.csv("tgw/2022/exports/to722.csv")
to922 <- read.csv( "tgw/2022/exports/to922.csv")
to1222 <- read.csv("tgw/2022/exports/to1222.csv")
code <- rbind(to418,to1018, to1619, to919, to919, to520, to621, to1121, to522, to722, to922, to1222)
code <- code[!duplicated(code),]

## Cleaining the raw data
drops <- c("IsDeleted","TotalFees","UnpaidFees","Value","CouncilDistrict","Next.Compliance.Date","CanPayFees","UserFavID","UserID","FavType","FavID","IsActive","IsDeleted1","ContactID","CanEdit","CanView","CanRenew","NeedAccept","Next.Hearing.Date","Last.Hearing.Date","LU_JurisdictionID","IsPublic","PubicDocCount","IsLockedExt", "IsLockedInt","Fav")
code <- code[, !(names(code) %in% drops)]
code <- mutate(code, FilingDate=as.Date(code$FilingDate, format="%m/%d/%Y %r"))
## We need to fix various errors in the data. This includes geocoding
## from addresses using https://www.gpsvisualizer.com/geocoding.html  

code[code$Location=="1552 Old Richmond Cir","Location"] <- "1252 Old Richmond Cir" 
code[code$Location=="1556 Old Richmond Cir","Location"] <- "1256 Old Richmond Cir"
code[code$Location=="0 Star Crest Dr","Location"] <- "1080 Starcrest Dr"
code[code$Location=="0 Suter St","Location"] <- "121 Suter St"
code[code$Location=="901 S High Street","Location"] <- "205 Sunrise Ave"
code[code$Location=="66 Pleasant Hill rd, 60 Pleasant Hill Rd","Location"] <- "66 Pleasant Hill Rd"

code[code$Location=="313 Old South High St", c("YWGS84")] <- c(38.44580084334057)
code[code$Location=="313 Old South High St", c("XWGS84")] <- c(-78.87584970859547)
code[code$Location=="205 Sunrise Ave", c("YWGS84")] <- c(38.441041699270514)
code[code$Location=="205 Sunrise Ave", c("XWGS84")] <- c(-78.88489286387966)
code[code$Location=="66 Pleasant Hill Rd", c("YWGS84")] <- c(38.428048249736136)
code[code$Location=="66 Pleasant Hill Rd", c("XWGS84")] <- c(-78.89348158194454)
code[code$Location=="1083 Lois Lane", c("YWGS84")] <- c(38.4165629683439)
code[code$Location=="1083 Lois Lane", c("XWGS84")] <- c(-78.87282929749068)
code[code$Location=="371 W Water St -Apt B", c("YWGS84")] <- c(38.449331296072266)
code[code$Location=="371 W Water St -Apt B", c("XWGS84")] <- c(-78.87555135613829)

missings <- unique(rbind(code[code$XWGS84> -20|code$XWGS84< -1000, c("Location","City") ],code[code$YWGS84> 2000|code$YWGS84< 30, c("Location","City")]))
missings$City <- "Harrisonburg"
missings$State <- "VA"
missings$Zip <- "22801"

write.csv(missings, file="missings.csv")
## Submit missings.csv to https://geocoding.geo.census.gov/geocoder/locations/addressbatch?form
## This will return GeocodeResults.csv to your downloads directory. Move it to your working directory and continue.
GeocodeResults <- read.csv("GeocodeResults.csv",  skip=1, header=FALSE)
GeocodeResults <- separate(GeocodeResults, col=c("V2"), sep=",", into=c("Location","City","State","Zip"), remove=TRUE, convert=TRUE)
GeocodeResults <- separate(GeocodeResults, col=c("V6"), sep=",", into=c("XWGS84","YWGS84" ), remove=TRUE, convert=TRUE)

for (i in 1:dim(missings)[1]){
code[code$Location==missings[i,"Location"], c("XWGS84")] <- GeocodeResults[GeocodeResults$Location==missings[i,"Location"], c("XWGS84")]
code[code$Location==missings[i,"Location"], c("YWGS84")] <- GeocodeResults[GeocodeResults$Location==missings[i,"Location"], c("YWGS84")]
}

code[code$Status=="","Status"] <- "No Status"
code[code$Status=="2nd Notice - Tall Weeds and Grass", "Status"] <- "2nd Notice - TGW"
code[code$Guid== "a4571e39-ab32-4e00-908b-da17f1237d60","Status"] <- "No Violations Found"

## Virginia FOIA law doesn't protect the names of people who make
## aesthetic nuisance complaints, but zoning inspectors are instructed
## to proactively mark such complaints as anonymous and withold their
## names. Building Code inspectors, on the other hand, make a note of
## the full name of tenants who make a building code complaint about
## a landlord. We proactively protect the identity of those tenants.

privacy <-c(1799,1794,1740,1741,1731,1718,1717,1692,1651,1634,1611,1593,1584,1583,1580,1560,1540,1537,1526,1523,1501,1431,1317,1258,1237,1195,1193,1189,1181,1180,1175,1167,1162,1149,1150,1143,1140,1139,1123,1107,1084,1082,1077,1063,1036,1030,1029,1026,1022,1018,1017,1016,1015,1014,998,993,989,975,969,965,957,955,951,935,922,917,916,892,893,884,883,877,871,870,869,867,864,862,860,859,848,845,841,819,762,728,724,718,705,704,703,701,693,692,668,666,636,630,628,584,551,531,487,467,465,455,450,407,391,390,375,287,281,236,234,224,211,199,196,176,166,132,122,47,46,27,7,1797,1775)

tenants <- grep("tenant",code$Descr, ignore.case=TRUE, value=FALSE)
privacy <- unique(c(tenants,privacy))
code[privacy,"Descr"] <- unlist(lapply(code[privacy,"Descr"], gsub, pattern="[0-9]+?", replacement="x"))
code <- code %>% filter( Guid!="661f722c-936e-4bd1-a974-eb37663b0a45"& Guid!="11b32ee3-b20f-49fa-87ac-621f9e00ab30")
```
We next read in historical tall grass and weed complaints which were
collected prior to the Department of Community Development switching
to its current data system, and clean the data so it can be merged.

```{r, early tgw, echo=FALSE, message=FALSE, cache=FALSE}
tgw12 <- read.csv("tgw2.csv")
tgw12 <- mutate(tgw12, Date.Received=as.Date(Date.Received))
tgw12 <- tgw12 %>% rename("Location"="Address")
tgw12 <- tgw12 %>% rename("Status"="Violation.Status.Open.Closed")
tgw12 <- tgw12 %>% rename("FilingDate"="Date.Received")
tgw12$Source.of.Concern <- "unknown"
levels(tgw12$Source.of.Concern) <- c("unknown","proactive")
tgw12$Source.of.Concern[c(130:133,122,120,114,115)] <- "proactive"

tgwraw13 <- read.csv("tgwraw13.csv")
tgwraw13 <- tgwraw13 %>% rename("Location"="Address")
tgwraw13 <- tgwraw13 %>% rename("Status"="Violation.Status.Open.Closed")
tgwraw13 <- mutate(tgwraw13, Date.Received=as.Date(tgwraw13$Date.Received,"%m-%d-%Y"))
tgwraw13 <- tgwraw13 %>% rename("FilingDate"="Date.Received")
## geocoded12 <- read.csv("addresses.csv")
## ## https://cran.r-project.org/web/packages/stringr/vignettes/regular-expressions.html
## geocoded12 <- separate(geocoded12, col="name", sep="\\sHarrisonburg", into=c("Location","name") )
geocoded13 <- read.csv("addresses13.csv")
geocoded13 <- separate(geocoded13, col="name", sep="\\sHarrisonburg", into=c("Location","name") )
tgw13 <- left_join(tgwraw13, geocoded13[,c("Location","latitude","longitude")], by="Location")
##tgw13 <- tgw13[tgw13$Source=="citizen concern",] # for natural yards, interested in staff inspection as well

earlytgw <- bind_rows(tgw12[,names(tgw13)],tgw13)

##names(earlytgw)[match("tgwfull$Comments",colnames(earlytgw))] <- "type"
##names(earlytgw)[match("Violation.Status.Open.Closed",colnames(earlytgw))] <- "Status"
earlytgw$TypeName <- "Tall Grass and Weeds"
earlytgw$type <- earlytgw$Comments
earlytgw$type <- as.factor(earlytgw$type)
levels(earlytgw$type)[grep("[mb]ul[tp][it][pi]le", levels(earlytgw$type))] <- "group"
levels(earlytgw$type)[grep("unk", levels(earlytgw$type))] <- "junk"
levels(earlytgw$type)[grep("inop", levels(earlytgw$type))] <- "car"
levels(earlytgw$type)[grep("oreclo", levels(earlytgw$type))] <- "foreclosure"
levels(earlytgw$type)[grep("bank owned", levels(earlytgw$type))] <- "foreclosure"
levels(earlytgw$type)[grep("asset company", levels(earlytgw$type))] <- "foreclosure"
levels(earlytgw$type)[grep("garden", levels(earlytgw$type))] <- "garden"
levels(earlytgw$type)[grep("back",levels(earlytgw$type))] <- "bounced"
levels(earlytgw$type)[grep("trouble finding",levels(earlytgw$type))] <- "bounced"
levels(earlytgw$type)[grep("[tT]all", levels(earlytgw$type))] <- "TGW"
levels(earlytgw$type)[grep("but", levels(earlytgw$type))] <- "TGW"
levels(earlytgw$type)[grep("RGR", levels(earlytgw$type))] <- "mowed"
levels(earlytgw$type)[c(24,10,9)] <- "mowed"
levels(earlytgw$type)[c(16,13)] <- "common"
levels(earlytgw$type)[c(18:21,15,16,10:12,6:8,4)] <- "TGW"


earlytgw$Status[grep("[0-9]", earlytgw$Status)] <- "No Status"
earlytgw$Status[grep("Closed", earlytgw$Status,ignore.case=TRUE)] <- "No Status" 
earlytgw$Status[grep("n-a", earlytgw$Status)] <- "No Status"
earlytgw$Status[earlytgw$Status=="no violation"] <- "No Violations Found"
earlytgw$Status[earlytgw$Status=="No violation"] <- "No Violations Found"
earlytgw$Status[earlytgw$Status=="unfounded"] <- "No Violations Found"
earlytgw$Status[earlytgw$Status==""] <- "No Status"
earlytgw$Status[earlytgw$Status=="Rosalyn"] <- "No Status"
earlytgw$Status[earlytgw$Status=="Generate Sign Civil Penalty Notice"] <- "Sign Notice"



## tgwdate <- earlytgw$FilingDate[earlytgw$Source.of.Concern=="unknown"|earlytgw$Source.of.Concern=="citizen concern"]
## earlytgw <- earlytgw[earlytgw$Source.of.Concern=="unknown"|earlytgw$Source.of.Concern=="citizen concern",]
breaksv <- as.Date(c("2011-04-01","2011-7-24","2011-09-01","2012-03-01","2012-07-24","2012-09-01","2012-12-1","2013-03-01","2013-07-24","2013-12-1"))
era <- cut(unclass(earlytgw$FilingDate),unclass(breaksv),lab=FALSE)
era <- as.factor(era)
levels(era) <- c("2011pre","2011post","2012pre","2012post","2012xtended","2013pre","2013post","2013extended")
era <- factor(era,levels=c("2011pre","2012pre","2013pre","2011post","2012post","2013post","2012xtended","2013extended"))
earlytgw$era <- era

earlytgw <- earlytgw %>% rename("Origin"="Source.of.Concern")
earlytgw <- earlytgw %>% rename("Descr"="Comments")
earlytgw <- earlytgw %>% rename("YWGS84"="latitude")
earlytgw <- earlytgw %>% rename("XWGS84"="longitude")
```

We next load demographic data and shape files [that we learned to pull from the Census Bureau](https://github.com/walkerke/umich-workshop-2022). 
Other sources include plotting [using google maps](https://rpubs.com/nickbearman/r-google-map-making) and [open street maps](https://ajsmit.github.io/Intro_R_Official/mapping-google.html).


```{r, demographicfs, eval=TRUE, echo=FALSE, message=FALSE}
options(tigris_use_cache = TRUE)

burglatino2020 <- get_decennial(year=2020, geography="block", county="Harrisonburg", state="VA", variables=c("P2_002N","P2_003N","P2_004N","P2_005N","P2_006N","P2_007N","P2_008N","P2_009N"), summary_var = "P2_001N",geometry=TRUE, cache_table=TRUE) %>% mutate(percent=100*(value/summary_value))

burgraceall2020 <- get_decennial(year=2020, geography="block", county="Harrisonburg", state="VA", variables=c("P1_002N","P1_003N","P1_004N","P1_005N","P1_006N","P1_007N","P1_008N","P1_009N"),summary_var="P1_001N",geometry=TRUE, cache_table=TRUE)

burgraceall2020 <- burgraceall2020 %>% mutate(percent=100*(value/summary_value))
tract_grass_sf <- filter(burgraceall2020, variable=="P1_003N")%>% mutate(white=value-filter(burglatino2020, variable=="P2_005N")$value) %>% mutate(percent=100*(value/summary_value)) #white

```

We transform the spatial coordinates so features line up correctly [using the simple features package]( https://r-spatial.github.io/sf/articles/sf5.html). 
We then use [Jonathan Duggins' tutorial](https://jonathanduggins.com/data-matters) for tidyverse functions such as filter.

```{r, spatial, echo=FALSE, message=FALSE}
 
crs_utm <- st_crs(burgraceall2020)
code_sp <- st_as_sf(code, coords = c("XWGS84", "YWGS84"), crs = crs_utm )
earlytgw_sp <- st_as_sf(earlytgw, coords = c("XWGS84", "YWGS84"), crs = crs_utm )

code_pz_sp <- filter(code_sp,Division=="P&Z") 
code_pz_nuisance_sp <- filter(code_pz_sp, TypeName %in% c("Junk, Trash and Debris","Tall Grass and Weeds","Inoperable Vehicle","Signage"))
code_pz_nuisance_sp <- bind_rows(code_pz_nuisance_sp, earlytgw_sp)
code_pz_nuisance_sp <- within(code_pz_nuisance_sp, TypeName <-factor(TypeName, levels=names(sort(table(TypeName),decreasing=TRUE))))
code_pz_nuisance_sp <- within(code_pz_nuisance_sp, Status <-factor(Status, levels=names(sort(table(Status),decreasing=TRUE))))

```

We next load icon fonts for plotting features onto our maps.

```{r, icons, echo=FALSE, message=FALSE }

fontawesome::fa_png("building-wheat", fill=cbbPalette["blue"])
system("mv building-wheat.png building-none.png")
fontawesome::fa_png("car-side", fill="grey")
fontawesome::fa_png("building-wheat")
fontawesome::fa_png("trash", fill=cbbPalette["yellow"])
fontawesome::fa_png("trash-can", fill=cbbPalette["yellow"])
fontawesome::fa_png("couch", fill=cbbPalette["red"])
building_wheat <- tmap_icons("building-wheat.png", height=12)
building_none <- tmap_icons("building-none.png", height=12)
car_side <- tmap_icons("car-side.png", height=12)
trash <- tmap_icons("trash.png", height=12)
trash_can <- tmap_icons("trash-can.png", height=12)
couch <- tmap_icons("couch.png", height=12)

```
More data cleaning to establish uniform violation variables. 
```{r, violation partitioning,  echo=FALSE, message=FALSE}
grass <- code_pz_nuisance_sp %>% filter(TypeName=="Tall Grass and Weeds"&!Status%in% c("No Violations Found","No Status","Inspection"))
nograss <- code_pz_nuisance_sp %>% filter(TypeName=="Tall Grass and Weeds"&Status%in% c("No Violations Found","No Status","Inspection"))
junk <- code_pz_nuisance_sp %>% filter(TypeName=="Junk, Trash and Debris")
couches <- junk[unlist(lapply(junk$Descr, str_detect, "ouch|urni|hair")),]
jnk <- junk[!unlist(lapply(junk$Descr, str_detect, "ouch|urni|hair")),]
nojnk <- jnk %>% filter(Status%in%c("No Violations Found","No Status","Inspection"))
jnk <- jnk %>% filter(!Status%in%c("No Violations Found","No Status","Inspection"))
cars <- code_pz_nuisance_sp %>% filter(TypeName=="Inoperable Vehicle")

write.csv(code_pz_nuisance_sp %>% filter(TypeName=="Tall Grass and Weeds"), "grasslist2.csv") # To be used for capture/recapture study.
st_write(code_pz_nuisance_sp, "grasslist2.shp", append=FALSE) # To be used for capture/recapture study.
```

We create the main map with the tmap package.
```{r, 'main map setup', eval = TRUE, echo=FALSE, message=FALSE}
tmap_mode("view")
demographic <- tm_shape(tract_grass_sf, name="Percent Non-Latino White") +
  tm_fill(
    col = "percent",
    palette = "-Greens",
    style = "cont",
    contrast = c(0.1, 1),
    title = "Percent Non-Latino White",
    id = "summary_value", #"GEOID",
    popup.vars = c("percent","value","summary_value"),
    showNA = FALSE,
    alpha = 0.8) +
    tm_borders(col = "darkgray", lwd = 0.7)

```

We add features with jitter to protect the privacy of people who have been targeted by complaints. To do as grassdot <- as_dot_density(value=) follow https://youtu.be/KMOTZij9qhU?t=9770 and tgw.R for tabling and left joining counts to polygons.  This video is also how you can use parcel shape files in the real estate information system to connect iNatuaralist coordinates to to addresses.

```{r, 'main map', echo=FALSE, message=FALSE , eval=TRUE}

grassmap <- demographic +
    tm_shape(grass, name="Tall Grass and Weed")+
    tm_symbols(shape=building_wheat, id = "Status",
               popup.vars = c("Origin") , size=.5, jitter=.3, alpha=.5)+
    tm_shape(nograss, name="Unfounded Tall Grass and Weed")+
    tm_symbols(shape=building_none, id = "Status",
             popup.vars = c("Descr","Origin") ,size=.5, jitter=.3, alpha=.5) +
    tm_shape(cars, name="Inoperable Vehicle")+
    tm_symbols(shape=car_side, id = "Origin",
             popup.vars = c("Descr","Status") , size=.5, jitter=.3) +
    tm_shape(jnk, name="Junk, Trash and Debris")  +
    tm_symbols( shape=trash, id = "Origin",
             popup.vars = c("Status"), size=.5, jitter=.3) +
    tm_shape(nojnk, name="Unfounded Junk, Trash and Debris")+
    tm_symbols(shape=trash_can, id = "Origin",
             popup.vars = c("Descr","Status") , size=.5, jitter=.3) +
    tm_shape(couches, name="Couch on Porch") +
    tm_symbols(shape=couch, id = "Origin",
               popup.vars = c("Status"), size=.5, jitter=.3, alpha=.5) + tm_layout("Aesthetic nuisance complaints in Harrisonburg Virginia October 2017 - July 2022 \n jittered with error for privacy." )

grassmap+ggtitle("Harrisonburg Aesthetic Nuisance Complaints", subtitle="Complaints 2011-2013, 2017-2022")
```

```{r, code analysis, echo=FALSE, message=FALSE}

tgw <- code_pz_nuisance_sp
tgw <- within(tgw, TypeName <-factor(TypeName, levels=names(sort(table(TypeName),decreasing=FALSE))))

```
## Code enforcement complaints over time.


```{r widegraphss, message=FALSE, echo=FALSE, fig.width=12, fig.height=4}

seasons <- as.Date(c("2011-4-1","2011-11-1","2012-4-1","2012-11-1","2013-4-1","2013-11-1","2014-4-1","2014-11-1","2015-4-1","2015-11-1","2016-4-1","2016-11-1","2017-4-1","2017-11-1","2018-4-1","2018-11-1","2019-4-1","2019-11-1","2020-4-1","2020-11-1","2021-4-1","2021-11-1","2022-4-1","2022-11-1"))

ggplot(data=tgw, aes(x=FilingDate, y=Status), col=TypeName)+ geom_jitter(alpha=.5, pch=1) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+scale_x_date( minor_breaks="year", date_labels="%y-%m", breaks= seasons)+ggtitle("All Planning and Zoning enforcement status")+xlab("TGW enforcement season")

ggplot(data=tgw, aes(x=as.Date(FilingDate,format="%m/%d/%Y %r"), fill=TypeName))+geom_histogram(binwidth=30, colour="white")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+scale_x_date( minor_breaks="year", date_labels="%y-%m" , breaks= seasons)+ggtitle("All Planning and Zoning enforcement status")+xlab("TGW enforcement season")
```

```{r widegraphss TGW, message=FALSE, echo=FALSE, fig.width=12, fig.height=4}
tgw <- tgw %>% filter(TypeName=="Tall Grass and Weeds")

status_l <- c("No Violations Found", "No Status",  "Inspection", "Verbal Notice", "Tall Weeds and Grass Notice", "Correction Notice", "Return Receipt Rc'd", "Door Hanger", "2nd Notice - TGW", "Violations Remedied")
status_l <- rev(status_l)

tgw <- within(tgw, Status <- factor(Status, levels= status_l ))

tgw %>%ggplot( aes(x=as.Date(FilingDate, format="%m/%d/%Y %r"), y=Status, col=Status))+ geom_jitter(alpha=.5, pch=1) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+scale_x_date( minor_breaks="year", date_labels="%y-%m" ,breaks= seasons)+ggtitle("TGW enforcement status")+xlab("TGW enforcement season")

tgw <- tgw %>% mutate(Status=relevel(Status, 8))

## tgw <- within(tgw, Status <-factor(Status, levels=names(sort(table(Status),decreasing=FALSE))))

ggplot(data=tgw, aes(x=as.Date(FilingDate,format="%m/%d/%Y %r"), fill=Status))+geom_histogram(binwidth=30, colour="white")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+scale_x_date( minor_breaks="year", date_labels="%y-%m" ,breaks=seasons)+ggtitle("TGW enforcement status")+xlab("TGW enforcement season")+geom_vline(xintercept=as.Date("2013-8-13"), col=cbbPalette["red"])+geom_text(label="MLK Renaming",aes(as.Date("2013-8-13"), 40))
```
	
## Sentiments toward natural yards in Harrisonburg

```{r, 'survey', echo=FALSE, message=FALSE}
nomow <- read.csv("nomow.csv")
nomow$supports[nomow$supports==" yes"] <- "yes"
nomow <- within(nomow, city <-factor(city , levels=names(sort(table(city),decreasing=TRUE))))
nomow[c(73,114),"different"] <- ""
with(nomow,table(supports))
```

On May 7, 2022, [WHSV](https://www.whsv.com/2022/05/06/no-mow-may-puts-lawnmower-pause-help-bees/) aired a story on No Mow May. On their [Facebook](https://www.facebook.com/WHSV.TV/posts/10159997517915336) page, they asked if viewers would consider participating in No Mow May. The `r sum(table(nomow$supports))` responses to the post were coded and analyzed. Notable features are that Harrisonburg overwhelmingly supported No Mow May with support of `r 100*(table(nomow$supports)["yes"]/sum(table(nomow$supports)))`%. Support in Harrisonburg featured acceptance of difference, whereas being different was viewed negatively outside Harrisonburg.

```{r, 'survey analysis', echo=FALSE, message=FALSE}
ggplot(data=nomow,aes(x=city, fill=supports))+ geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("WHSV May 7 Facebook responses on No Mow May preference", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")

ggplot(data=nomow %>% subset(supports=="yes"), aes(x=city, fill=different))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("Difference theme by No Mow May supporters.", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
ggplot(data=nomow %>% subset(supports=="no"), aes(x=city, fill=different))+geom_bar()+theme(axis.text.x = element_text(an gle = 90, vjust = 0.5, hjust=1))+ggtitle("Difference theme by No Mow May opponents.", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")


ggplot(data=nomow %>% subset(supports=="yes"), aes(x=city, fill=HOA))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("HOA theme by No Mow May supporters.", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
ggplot(data=nomow %>% subset(supports=="no"), aes(x=city, fill=HOA))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("HOA theme by No Mow May opponents.", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")



ggplot(data=nomow %>% subset(supports=="yes"), aes(x=city, fill=laws))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("Laws theme by No Mow May supporters", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
ggplot(data=nomow %>% subset(supports=="no"), aes(x=city, fill=laws))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("Laws theme by No Mow May opponents", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")


ggplot(data=nomow %>% subset(supports=="yes"), aes(x=city, fill=mowing))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("Difficult to mow theme by No Mow May supporters", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
ggplot(data=nomow %>% subset(supports=="no"), aes(x=city, fill=mowing))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("Difficult to mow theme by No Mow May opponents", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")


ggplot(data=nomow %>% subset(supports=="yes"), aes(x=city, fill=critters))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("Attract critters theme by No Mow May supporters", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
ggplot(data=nomow %>% subset(supports=="no"), aes(x=city, fill=critters))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("Attract critters theme by No Mow May opponents", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")

ggplot(data=nomow %>% subset(supports=="yes"), aes(x=city, fill=flowers))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("Have flowers theme by No Mow May supporters", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
ggplot(data=nomow %>% subset(supports=="no"), aes(x=city, fill=flowers))+geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("Have flowers theme by No Mow May opponents", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")


harrisonburg <- nomow %>% subset(city=="Harrisonburg")
ggplot(data=harrisonburg, aes(x=supports, fill=different))+geom_bar() +ggtitle("Difference theme by No Mow May Harrisonburg residents", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
outsiders <- nomow %>% subset(city!="Harrisonburg")
ggplot(data=outsiders, aes(x=supports, fill=different))+geom_bar() +ggtitle("Difference theme by No Mow May non-Harrisonburg residents", subtitle="https://www.facebook.com/WHSV.TV/posts/10159997517915336")
```
	
